Lazy Linearity for a Core Functional Language

### 2025-10-20 popl26.tex
added coverpage for supplement
Move PRINCIPLE from S1 to S3
Delete 1 backref from related work

Linear Mini-Core~\cite{cite:minicore} is a specification of linear types in
Core as they were being implemented in GHC, and doubles as the (unpublished)
precursor to our work. Linear Mini-Core first observes the incapacity of
Core's type system to accept linear programs after transformations, and 
introduces usage environments for let-bound variables with the same goal of
Linear Core of specifying a linear type system for Core that accepts the
optimising transformations.
%
We draw from Linear Mini-Core the rule for non-recursive let expressions and
how let-bound variables are annotated with a usage environment. However, our
work further explores the interaction of laziness with linearity in depth, and
diverges significantly in rules for typing other constructs, notably, case expressions and
case alternatives. Furthermore, unlike Mini-Core, we prove Linear Core is type
safe, guarantees linear resource usage, and that multiple optimising
transformations, when applied to Linear Core programs, preserve linearity.
\ROUNDTWO{As previously noted in the paper, our irrelevant resources
  are similar to 0-multiplicity
  values from QTT~\cite{brady:LIPIcs.ECOOP.2021.9} and our tagged resources are
  related to fractional permissions from separation
  logic~\cite{10.5555/1760267.1760273}.}

@@

>>>>

Linear Mini-Core~\cite{cite:minicore} is a specification of linear types in
Core as they were being implemented in GHC, and doubles as the (unpublished)
precursor to our work. Linear Mini-Core first observes the incapacity of
Core's type system to accept linear programs after transformations, and 
introduces usage environments for let-bound variables with the same goal of
Linear Core of specifying a linear type system for Core that accepts the
optimising transformations.
%
We draw from Linear Mini-Core the rule for non-recursive let expressions and
how let-bound variables are annotated with a usage environment. However, our
work further explores the interaction of laziness with linearity in depth, and
diverges significantly in rules for typing other constructs, notably, case expressions and
case alternatives. Furthermore, unlike Mini-Core, we prove Linear Core is type
safe, guarantees linear resource usage, and that multiple optimising
transformations, when applied to Linear Core programs, preserve linearity.
\ROUNDTWO{Our irrelevant resources
  have similarities with 0-multiplicity
  values from QTT~\cite{brady:LIPIcs.ECOOP.2021.9} and, as previously noted,
  our tagged resources are related to fractional permissions from separation
  logic~\cite{10.5555/1760267.1760273}.}

@@

\newpage

~\vspace{8cm}
\begin{center}
  {\huge
    Supplemental Materials for \\ Lazy Linearity for a Core Functional
    Language 
    }
\end{center}

\newpage
